\section{Existing Works}
Due to the large number of network intrusion detection systems that adopt machine learning or data mining approaches
we only review some of them that achieved state-of-the-art detection performance.
There are very limited number of existing network intrusion detection systems that adopt
deep learning approaches.
Their details can be found in section~\ref{Sec:Architectures}.

\subsection{State-of-Art Machine Learning Approaches}
Prior researchers modeled the intrusion detection task as an unsupervised
anomaly detection problem, and proposed a series of approaches.
Examples include Mahalanobis-distance based outliner detection~\cite{ComparativeAnomalyNIDS},
density-based outliner detection~\cite{LOF, ComparativeAnomalyNIDS},
evidence accumulation for ranking outliner~\cite{RankingOutliner}, etc.
One of the advantage of these unsupervised approaches is to tackle the problem of
the unavailability of labeled traffic data.

Alternatively, prior researchers made a lot of effort to obtain meaningful
attacking data and to convert them into large amount of labeled data~\cite{DARPA, KDDCup, NSL-KDD, UNSW, UNSW1}.
Such efforts make it possible to apply supervised machine learning algorithms to the
intrusion detection problem.
Successfully applied approaches include decision trees~\cite{DecisionTree},
linear and non-linear support vector machines~\cite{SVM}, NB-Tree~\cite{NB-Tree} and so on.
To the best of the authors' knowledge, there are two works that achieved the best prediction
accuracy on two different dataset respectively.
For the UNSW-NB15 dataset, it is reported in~\cite{RampLossKSVCR} that extending K-support vector
classification-regression~\cite{KSVCR} with ramp loss, called Ramp-KSVCR approach, can achieve an accuracy of 93.52\%.
The authors of Ramp-KSVCR also report that their approach can achieve 98.68\% accuracy on the NSL-KDD dataset.
On the other hand, the creators of UNSW-NB15 dataset~\cite{UNSW} proposed an approach called
Geometric Area Analysis techniques using trapezoidal area estimation (GAA-ADS for short)~\cite{GAA-ADS}.
It achieves even better accuracy on NSL-KDD dataset (99.7\%) and a slightly worse accuracy on
UNSW-NB15 dataset (92.8\%).


\subsection{Deep Learning Flavor Approaches}
There are some pioneer works that introduced deep learning approaches to intrusion detection.
For example, \cite{STL-NIDS} adopts sparse autoencoder and the self-taught learning
scheme~\cite{SparseAE} to handle the problem that there is limited amount of labeled data
available for training supervised model.
Similar semi-supervised approach have also been applied to
Discriminative restricted Boltzmann machine~\cite{AnomalyDetectionRBM}.
